A Data-Driven Framework for Neural Field Modeling

### 2010-11-21 ltx/first_order_framework/LaTeX_Manuscript.tex
minor correction in appendixB added bracket to formula B1
minor correction in frequency analysis
minor correction in table3

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. Again, applying Shannon's sampling theorem, the minimum distance between basis functions to accurately reconstruct the neural field must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{cy}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

@@

	$\sigma_{\phi}$ & width of field basis functions in spatial domain & 2.5 & mm$^2$\\

@@

\section{Observation Kernel}\label{App:ObsKernelExperiment}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Fig.~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a center frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{T}\sum_n{\sqrt{(y_t-\mathcal{H}y_t^2}}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $t$ indexes samples and $T$ is the total number of samples in the time series. Fig.~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution.
\section{Inner Product of Two \emph{n}-Dimensional Gaussians}\label{App:InnerProdOfGaussians}
The derivation in this section was used for calculating the matrix $\mathbf \Gamma$ from equation~\ref{eq:DefGamma}. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}

>>>>

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. Again, applying Shannon's sampling theorem, the maximum distance between basis functions to accurately reconstruct the neural field must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{cy}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

@@

	$\sigma_{\phi}$ & width of field basis functions in spatial domain & 1.58 & mm$^2$\\

@@

\section{Observation Kernel}\label{App:ObsKernelExperiment}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Fig.~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a center frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{T}\sum_n{\sqrt{(y_t-\mathcal{H}y_t^2})}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $t$ indexes samples and $T$ is the total number of samples in the time series. Fig.~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution.
\section{Inner Product of Two \emph{n}-Dimensional Gaussians}\label{App:InnerProdOfGaussians}
The derivation in this section was used for calculating the matrix $\mathbf \Gamma$ from equation~\ref{eq:DefGamma}. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}
